Machine Learning Methods for Discourse Segmentation of Communications in E-Mail Based Behavioral Interventions

### 2018-07-30 paper-amia-summit-2018/manuscript.tex
update paper

Qualitative coding of motivational interviews with pre-defined codes has been traditionally performed manually by trained annotators, which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts. Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process. In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis\cite{wang2012baselines}. In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and analysis \cite{hasan2018predicting} of in-person MI sessions. Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders\cite{hasan2016study}. Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient and a caregiver.

@@

The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis. A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors. Automating this task is a unique and challenging problem due to the following major reasons:

@@

%Our experiments show that these differences lead to significant improvement in the segmentation of e-Coaching emails.

Recently, an online clinical intervention called MENU GenY \cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated. MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30. The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits. However, segmentation of clinical conversation in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions. This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.

@@

The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word or punctuation marks annotated with a label 1 or 0, to indicate whether it precedes a new segment or not. In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN. For MLP models, samples were created with $2l$ words/punctuation marks at each step in a word sequence, where a sample contains next $l$ words/punctuation marks and prior $l$ words/punctuation marks including current word/punctuation mark. Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word/punctuation mark precedes a new segment or not. For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation marks were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used topic features for our experiment \cite{liu2005using,treviso2017sentence}. To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary. For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}. Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words. Trained word vectors were also utilized when punctuation marks are not available in the pre-trained word2vec word vectors. Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/embedding-dimension.eps}
    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.}}
    \label{fig:embedding-dimension}
\end{figure}

@@

In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three neural networks models. Since neural networks provide a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.

@@

Here $Z_x$ is a normalization factor over all possible labelings of $x$, and $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$. The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm. In our experiments, the following features were utilized with CRF models: i) current word/punctuation ii) next and previous 3 words/punctuations iii) whether the word/punctuation is special characters (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is title or not? (e.g. "The" is a title word but "the" is not) v) POS tags

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural network model which maps set of input onto a set of appropriate output. This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase \cite{srivastava2014dropout}. Dropout is also applied in the fully connected layers in BRNN and CRNN models. For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors. Figure~\ref{fig:length-mlp} demostrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.60\textwidth]{figures/length-mlp.eps}
    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size.}}
    \label{fig:length-mlp}
\end{figure}

\textbf{Bidirectional Recurrent Neural Networks}: BRNN is a neural network designed to capture sequential patterns by considering both past and future inputs as well as complex relationships between input feature and output labels. The output of BRNN layer is computed as the summation of the forward RNN output with backward RNN output. Gated Recurrent Units (GRU) \cite{chung2014empirical} are a special type of RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. Figure~\ref{fig:crnn} will represent an architecture of BRNN if convolution layer is removed.

@@

A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After the convolutional layer, $m \times n_f$ inputs are fed into bidirectional recurrent neural networks with 200 hidden units. Dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word precedes a new segment or not. We experimentally determined that the best performance is achieved when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. Source codes of all experiments will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection. CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes. BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure. On the other hand, MLP has the highest precision of 0.836 when only word embeddings or lexical features are used to identify ``new segment''. CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes. Experimental results indicate that performance of classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified. For example, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN shows the lowest and CRF shows the second highest performance for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes.\\

@@

\section*{Discussion}
This study is the first effort to evaluate the automatic segmentation of e-Coaching emails. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''. The robust performance of CRNN provides the evidence that machine learning models are capable to learn conceptual information from the clinical exchange. It also indicates that punctuation and POS features are more important than word embeddings or lexical features, even when deep learning methods are employed. Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings/lexical (left) and combination of word embeddings/lexical, POS and punctuation features (right) are used}.}
    \label{fig:roc-curves}
\end{figure}

In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance. In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets. AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively. \\

\begin{table}[ht]
\centering
\caption{AUC values of all classifiers when only word embeddings/lexical and combination of word embeddings/lexical, POS and punctuation features are used. Highest value for each experiment across all models is highlighted in boldface.}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{Area Under the ROC Curve (AUC)}} \\\cline{2-5}
 & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings or lexical features & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

@@

Although punctuation mark plays an important role in segmentation boundary detection, large numbers of errors were encountered by the false positive of boundary identification. For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack. Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times. I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I quotemark m hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered. Similarly, additional information is a common cause for misclassification of an email segment into multiple segments. For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.

@@

\section*{Conclusion}
Segmentation of e-Coaching emails is an integral part of developing e-Coaching interventions. Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions. In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails. In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text. We found out that CRNN provides the best performance for the segmentation of text in terms of all performance metrics. Manual segmentation of e-Coaching data is very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention. Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model. This approach will also help for developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.

>>>>

Qualitative coding of motivational interviews with pre-defined codes, has been traditionally performed manually by trained annotators which is a tedious and resource-intensive process that involves several iterations of reading, comprehension and interpretation of interview transcripts. Rapidly developing computational technologies, specifically, machine learning methods, offer a unique opportunity to accelerate this process. In particular, machine learning methods have been successfully applied to a variety of analytical tasks involving textual data, such as classification\cite{nigam2000text} and sentiment analysis\cite{wang2012baselines}. In our previous work, we examined the utility of machine learning methods for automated annotation \cite{hasan2016study,kotov2015interpretable} and analysis \cite{hasan2018predicting} of in-person MI sessions. Specifically, we demonstrated that machine learning methods can be utilized for annotation of MI transcripts according to a simple communication code scheme with the accuracy comparable to human coders\cite{hasan2016study}. Experimental data utilized in these studies, however, were prepared by transcribing audio conversations, which were clearly segmented into utterances by a counselor, a patient and a caregiver.

@@

The unstructured nature of e-Coaching exchanges poses a unique set of challenges for their qualitative analysis. Usually, email segments are automatically code with the Minority Youth-Sequential Coding of Process Exchanges (MYSCOPE \cite{carcone2013provider}), a qualitative coding scheme to characterize client-coach communication during email exchange. A significant barrier to fully automating the behavior coding process of e-Coaching emails is their segmentation into textual fragments that correspond to distinct communication behaviors. Automating this task is a unique and challenging problem due to the following major reasons:

@@

Recently, an online clinical intervention called MENU GenY \cite{alexander2017motivations} (Making Effective Nutrition Choices for Generation Y) was proposed and evaluated. MENU GenY is a technology-based public health intervention that relies on personalized e-coaching to encourage increased fruit and vegetable intake among young adults, aged 21-30. The goal of MENU GenY was to develop a better coding dictionary among GenY to improve eating habits. However, segmentation of clinical exchange in the context of electronically delivered interventions, in particular, segmentation of clinical interaction text into groups of MI behaviors, is still performed manually, which slows down qualitative analysis of these interventions. This study proposes several machine learning methods to address this problem and the authors are unaware of any other work that focused on the same problem.

@@

The experimental dataset for this work was constructed from 49 e-coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We formulate the segmentation task as a binary classification problem where each word or punctuation mark annotated with a label 1 or 0, to indicate whether it precedes a new segment or not. In total, we obtained 95,777 words and 7,140 punctuation marks, which include 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning method, CRF and deep learning methods, such as MLP, BRNN and CRNN. For MLP models, samples were created with $2l$ words/punctuations at each step in a word sequence, where a sample contains next $l$ words/punctuations and prior $l$ words/punctuations including current word/punctuation. Each sample is classified into either ``new segment'' or ``same segment'' based on whether the current word/punctuation precedes a new segment or not. For CRF, BRNN and CRNN models, an email was taken as input sequence, such that POS tags and word embedding of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decision were considered as the output of the model at each step. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

We utilized three types of features in conjunction with CRF, MLP, BRNN and CRNN methods: word embeddings or lexical features, punctuation and POS features. Since POS tags have been shown to be effective semantic abstractions of individual words, we used POS features for our experiment \cite{liu2005using,treviso2017sentence}. To extract our POS features, we tag the e-Coaching emails using the NLTK POS tagger \cite{bird2009natural}. Punctuation marks, which correspond to one of the symbols \{`.', `,', `!', `?', `:', `;'\} between a pair of words, are also employed as a feature since punctuation marks designate the boundary of a sentence, clause and phrase and often also correspond to a segment boundary. For natural language processing (NLP) tasks, inputs are received as a text where words can be considered as the basic processing unit. Therefore, it is important to represent a word in such a way that it carries all relevant information. Word embeddings are one of this representation where each word represented as a real-valued vector in a high dimensional vector space. The distributed representation of word allows capturing semantic, syntactic and morphological information from large unannotated corpora  \cite{pennington2014glove, mikolov2013distributed}. Since pre-trained word2vec word vectors provide the best results, in our experiment, we utilized pre-trained word2vec word vectors for words. Trained word vectors were also utilized when rare words or punctuation marks are not available in the pre-trained word2vec word vectors. Figure~\ref{fig:embedding-dimension} illustrates the performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models. It was observed that best performance achieved with pre-trained 300 dimensional word2Vec word vectors when three types of features are used together. Therefore, we report our results with word2vec 300 dimensional word vectors for all deep learning models used in this study. CRF utilized lexical features which represent each word in an e-Coaching email.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/embedding-dimension.eps}
    \caption{\textbf{Performance of CRNN model on e-Coaching email segmentation by varying the dimension of pre-trained and trained word embeddings with GloVe and word2vec models.}}
    \label{fig:embedding-dimension}
\end{figure}

@@

In the present study, we experimented with four different classifiers, including one traditional machine learning method, CRF and three deep learning models. Since deep learning provides a flexible architecture for constructing and synthesizing complex models, we take the advantage of this flexibility to test MLP, BRNN and CRNN models for the task of segmentation of e-Coaching emails.

@@

Here $Z_x$ is a normalization factor over all possible labelings of $x$, and $f_k(y_{t-1}, y_t, x, t)$ is a feature function, and $\lambda_k$ is a learned weight associated with feature $f_k$. The optimal output sequence $y^*$ for an input sequence $x$, $y^* = {arg\,max}_y p(y|x)$, is obtained efficiently by the Viterbi algorithm. In our experiments, the following features were utilized with CRF models: i) current word/punctuation ii) next and previous 3 words/punctuations iii) whether the word/punctuation is special characters (';', '?', '.', ',', '!', ':', etc) or not iv) whether the word is title word or not? (e.g. ``The'' is a title word but ``the'' is not) v) POS tags

\textbf{Multilayer Perceptrons}: MLP is a feed-forward artificial neural networks which map set of input onto a set of appropriate output. This study utilized MLP with a nonlinear activation function (relu) and one hidden layer consisting of 128 hidden units. In order to prevent over-fitting, we implement dropout on fully connected layers, which randomly hide some neurons during the training phase \cite{srivastava2014dropout}. Dropout is also applied in the fully connected layers in BRNN and CRNN models. For a given sample with $2l$ words/punctuations, the first layer input of MLP was created by the summation of first $l$ word vectors and last $l$ word vectors. Figure~\ref{fig:length-mlp} demostrates the performance of MLP model on e-Coaching email segmentation by varying value of $l$. Based on the results in Figure~\ref{fig:length-mlp}, MLP results were reported with $l = 2$, since best performance of MLP models were observed when $l = 2$.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/length-mlp.eps}
    \caption{\textbf{Performance of MLP model on e-Coaching email segmentation by varying the length of sample size.}}
    \label{fig:length-mlp}
\end{figure}

\textbf{Bidirectional Recurrent Neural Networks}: BRNN is a neural network designed to capture sequential patterns by considering both past and future inputs as well as complex relationships between input feature and output labels. The output of BRNN layer is computed as the summation of the forward RNN output with backward RNN output. Gated Recurrent Units (GRU) \cite{chung2014empirical} are utilized as RNN that are capable of handling variable size input sequence and have an internal memory that can be reset. Figure~\ref{fig:crnn} will represent an architecture of BRNN if convolution layer is removed.

@@

A sequence of $m$ words is fed into the input layer which provides a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embedding, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = n_w + n_p$ of size 310. The 1D convolutional layer generates $m - h_c + 1$ new features with $h_{c.n_e}$, a filter of $h_c$ words and zero-padding on both sides of the text. A pooling operation was performed over time to find the most significant features in a region. After applying $n_f=100$ different filters on convolutional layer, $m \times n_f$ inputs are generated and fed into bidirectional recurrent neural networks, which contain 200 hidden units. Then, dropout is used to randomly drop 50\% of the neurons and passes them through a fully connected layer, where a sigmoid operation is calculated, giving us the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined that the best performance is achieved with 5 folds corss-validation when filter length in convolution layer is 7, pool size is 3, batch size is 32, adam\cite{kingma2014adam} is used for optimization and the early stopping strategy is used with dropout to prevent over-fitting. To reuse of our source codes for other research studies, source codes of all experiments will be publicly available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication.

@@

As follows from Table~\ref{tab:result_base}, CRNN outperforms all other models in terms of recall and F1-measure achieving 0.797 recall with 0.785 F1-measure for new segment detection. CRNN also shows superior performance in all performance metrics for weighted average over ``new segment'' and ``same segment'' classes. BRNN demonstrates the lowest performance among all models in terms of precision and F1-Measure. On the other hand, MLP has the highest precision of 0.836 when word embeddings or lexical features are used to identify ``new segment''. CRF achieves 0.733 F1-Measure in new segment class and 0.984 F1-Measure in overall, which demonstrates the second highest performance to identify ``new segment'' as well as a weighted average over both classes. Experimental results indicate that performance of all classifiers as a weighted average over both classes is remarkably higher compared to ``new segment'' class, which is expected since 96.95\% instances belong to the ``same segment'' categories and 99.3\% of them correctly classified. For example, as a weighted average over ``new segment'' and ``same segment'' classes, CRNN achieves 27.23\%, 23.71\% and 25.61\% higher precision, recall and F1-measure, respectively, compared to ``new segment'' detection.

Table~\ref{tab:result_weighted_avg} summarizes the results of all models for segmentation of e-Coaching emails when word embeddings or lexical features are used in combination with punctuation and POS features. Similar to results in Tables~\ref{tab:result_base}, CRNN demonstrates the highest performance among all methods achieving 0.864 recall with 0.825 F1-Measure for ``new segment'' class and 0.990 precision with 0.989 recall and F1-Measure for overall. BRNN and CRF show the lowest and second highest performance, respectively, for email segmentation among all methods. We observed that classification performance significantly improved for ``new segment'' class when word embeddings or lexical features are used in combination with punctuation and POS features. Precision increases by 3.96\%, -2.63\%, 12.71\% and 1.81\%; recall increases by 11.72\%, 23.44\%, 20.59\% and 8.41\%; and F1-measure increases by 8.05\%, 11.1\%, 16.22\% and 5.1\% for CRF, MLP, BRNN and CRNN methods, respectively, in new segment detection. Similarly, precision increases by 0.51\%, 0.51\%, 0.82\% and 0.41\%; recall increases by 0.41\%, 0.41\%, 0.72\% and 0.3\%; and F1-measure increases by 0.41\%, 0.51\%, 0.82\% and 0.3\% for CRF, MLP, BRNN and CRNN methods, respectively, in weighted average over ``new segment'' and ``same segment'' classes.\\

@@

\section*{Discussion}
This study is the first effort to evaluate the automatic segmentation of e-Coaching emails. Experimental results indicate that CRNN is the best model among all machine learning methods considered for this study. CRNN achieved 0.989 F1-measure in overall and 0.825 F1-measure for detecting ``new segment''. The robust performance of CRNN provides the evidence that deep learning models are capable to learn MI behaviors from the clinical exchange. It also indicates that punctuation and POS features are important with word embeddings or lexical features, even when deep learning methods are employed. Although the domain of this study was intentionally quite small, we believe that our study is not limited to the e-Coaching domain and it can be successfully applied to other domains, in which discourse segmentation is a preliminary step for annotation.

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=1.0\textwidth]{figures/roc-curves.eps}
    \caption{\textbf{Receiver operating characteristic curves showing the performance of binary classifiers for the segmentation of e-Coaching text when only word embeddings/lexical features (left) and combination of word embeddings/lexical, punctuation and POS features (right) are used}.}
    \label{fig:roc-curves}
\end{figure}

\begin{table}[ht]
\centering
\caption{AUC values of all classifiers when only word embeddings/lexical features and combination of word embeddings/lexical, punctuation and POS features are used. Highest value for each experiment across all models is highlighted in boldface.}
\label{tab:result_roc}
 \begin{tabular}{|l|l|l|l|l|}
  \hline
\multirow{2}{*}{\textbf{Features}} & \multicolumn{4}{|c|}{\textbf{Area Under the ROC Curve (AUC)}} \\\cline{2-5}
 & \textbf{CRF} & \textbf{MLP}  & \textbf{BRNN} & \textbf{CRNN} \\ \hline      
 word embeddings or lexical features & \textbf{0.966} & 0.951 & 0.920 & 0.965 \\ \hline
 all features & \textbf{0.994} & 0.992 & 0.981 & 0.986 \\ \hline
  \end{tabular}
\end{table}

@@

In this paper, results are reported by ``new segment'' class as well as weighted average over both classes to avoid confusion regarding overall model performance due to severe class imbalance. In addition, standard metrics: precision, recall and F1-measure were used to eliminate doubt about the model performance since accuracy is misleading for imbalanced datasets. AUC is also utilized due to its effectiveness in measuring the quality of binary classifiers for imbalanced datasets\cite{hu2015kernelized}, which are demonstrated by the Table~\ref{tab:result_roc} and ROC curves in Figure~\ref{fig:roc-curves}. Although CRF provides the second highest classification results, it shows best AUC values, achieving 0.966 AUC when only lexical features are used and 0.994 AUC when a combination of lexical, punctuation and POS features are used. Influence of the punctuation and POS features is also consistent in AUC values, which increase by 2.9\%, 4.31\%, 6.63\% and 2.18\% for CRF, MLP, BRNN and CRNN methods, respectively.

Although punctuation mark plays an important role in segmentation boundary detection, a few errors were encountered by the presence of punctuation marks in boundary identification. For example, a text segment from an e-Coaching email \textit{``A typical day in regards to fruit and vegetable has me eating about a serving at breakfast (our caf has cut up fruit) and then maybe a piece of fruit later in the day or as a snack. Vegetable tends to be a side serving at lunch and dinner and I get celery or carrot cuts with dressing for a snack a lot of times. I could probably add some sort of vegetable into my breakfast (like spinach in an omelet) and snack on another piece of fruit when I quotemark m hungry rather than the junk food I tend to eat.''} incorrectly segmented after the first sentence where a punctuation mark was encountered. Similarly, additional information is a common cause for misclassification of an email segment into multiple segments. For instance, the first sentence of the above email segment represents a positive commitment to adolescent's behavior change then next two sentences provide an additional information to support their commitment.

@@

\section*{Conclusion}
Segmentation of e-Coaching emails is an integral part of developing e-Coaching interventions. Although several studies have focused on clinical interventions, they are limited by the qualitative coding of clinical interactions. In addition, previous studies in the medical domain mainly segmented clinical text into sections and sentences, none of them considered segmentation of text into groups of MI behaviors in the setting of discourse analysis with emails. In this paper, we compared the performance of machine learning models for the task of segmentation of e-Coaching text. We found out that CRNN provides the best performance for the segmentation of text in terms of all performance metrics. Manual segmentation of e-Coaching data is a very resource-intensive and time-consuming task, which can significantly decrease the time and effort required to develop an effective behavioral intervention. Our proposed methods can help to identify individual text segments, which can be annotated directly with a classification model. This approach will also help for developing fully automated e-Coaching and accelerate the pace of identifying effective communication strategies.
